---
title: "H2O Automl BOSTON"
author: "DA"
date: "March 10, 2018"
revised: "February 26, 2020"
---

# Install Packages

```{r}
# Remove any previously installed H2O packages for R
if ("package:h2o" %in% search()) { detach("package:h2o", unload=TRUE) }
if ("h2o" %in% rownames(installed.packages())) { remove.packages("h2o") }

# Download dependencies
pkgs <- c("RCurl","jsonlite")
for (pkg in pkgs) {
if (! (pkg %in% rownames(installed.packages()))) { install.packages(pkg) }
}

# Install the latest open source H2O package (see https://www.h2o.ai/download/#h2o)
install.packages("h2o", type="source", repos="http://h2o-release.s3.amazonaws.com/h2o/master/4970/R")
```

```{r}
# Install package for sample datasets
install.packages("mlbench")
```

# Load Packages

```{r}
# Libraries
library(h2o) # H2O Machine Learning
library(mlbench) # Datasets

n_seed = 42 # seed for reproducibility and everything...
```

# Load and Check Data

```{r}
# load dataset from mlbench and display its structure
data("BostonHousing")
bh = as.data.frame(BostonHousing) # simple, clean, classic dataset for regression 
str(bh) # FACTOR = CATEGORICAL
```

```{r}
# view a sample of the dataset

tail(bh)
```

```{r}
# check for missing values

unlist(lapply(bh, function(x) any(is.na(x)))) 
```

```{r}
# check for duplicate values

which(duplicated(bh)) 
```

## Define the target to predict and the feature columns to use

```{r}
target = "medv" # Median House Value
features = setdiff(colnames(bh), target)
print(features)
```

## Initialise h2o

```{r}
# Start a local H2O cluster (JVM compute engine)

h2o.init()
h2o.no_progress() # disable progress bar for RMarkdown
h2o.removeAll()   # delete objects from previous session (optional)
```

## convert R DF to h2o DF

```{r}
# H2O dataframe

h2o_boston = as.h2o(bh)
head(h2o_boston)
```

```{r}
# Note H2O DF structure

str(h2o_boston)
```


```{r}
# Split Train/Test

h_split = h2o.splitFrame(h2o_boston, ratios = 0.8, seed = n_seed)
h_train = h_split[[1]] # 80% for modelling
h_test = h_split[[2]] # 20% for evaluation
```

## AutoML - Train multiple models and a stacked ensemble

```{r}
# Train multiple H2O models with H2O AutoML
# H2O DL algorithm on multi-core is stochastic

model_automl = h2o.automl(x = features,
                          y = target,
                          training_frame = h_train,
                          nfolds = 5,               # Cross-Validation
                          max_runtime_secs = 300,   # Max time (5 mins)
                          max_models = 100,         # Max no. of models
                          stopping_metric = "RMSE", # Metric to optimize
                          project_name = "automl_reg",
                          exclude_algos = NULL,     # If you want to exclude any algorithms 
                          seed = n_seed)
```

## View AutoML Leaderboard

```{r}
model_automl@leaderboard
```

## Best Performing Model

```{r}
# Best Model (from an individual algorithm or a stacked ensemble)

model_automl@leader
```

## Evaluate Performance on unseen data

```{r}
# Remember lower value error metrics are better!

h2o.performance(model_automl@leader, newdata = h_test) 
```

## Make Predictions 

```{r}
ypred_test = h2o.predict(model_automl@leader, h_test)
head(ypred_test)
```

## Export Models 

- `h2o.saveModel()` to persist model
- `h2o.loadModel()` to re-load model
- `h2o.download_mojo()` for model serving

```{r, eval=FALSE}
# Save model to disk

h2o.saveModel(object = model_automl@leader, 
              path = "./",
              force = TRUE)
```

```{r}
sessionInfo()
```

